Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2016-02-02 GECCO/volunteer-clouds-gecco16.tex
Revised the intro and rewritten in part.
Remember that now we will only have the experiments, so we must tell as much about the server as possible in the intro. Hope it is enough. Please @mariosky @unintendedbear have a look.

Using the browser as a computing resource in volunteer computing presents several advantages, 
but it remains a challenge to harness the browser's capabilities and
to model the user's behavior so that those capabilities can be
leveraged optimally. In this paper, we present the results of several
experiments with a volunteer computing framework called {\sf NodIO}
that is easily deployable in using free cloud resources. We use 
%Pablo: "deployable in using free cloud resources" is correct?
% why not? - JJ
different implementations to find out which one is 
able to get the user to lend more computing cycles and test different
problems to check the influence it has on said performance.
% Pablo: so the objective is "to determine which browser conf. is able
% to get more computing power from the users", it isn't? 
% (to keep this in mind the rest of the paper). What is a "browser configuration"?
% (explain it with a phrase in the abstract)
% No, we don't. Please read the paper!!!! - JJ
From these experiments we can already draw some conclusions, besides
the fact that volunteer computing can be a valuable computing
resource: the user has to be kept engaged in some way to be able to
use as many computing cycles as possible, the client has to be built
to use the computer capabilities fully, and, finally, the user
contributions follow a common statistical distribution. It is also
essential to be as open as possible with software and data.

@@

Nowadays there is a big and increasing number of
distributed evolutionary computation software frameworks available on
the market, usually as open source frameworks \cite{Parejo12Survey}. 
Most of these frameworks have 
two features: they are {\em desktop} applications, that is, they must
be compiled or run directly on the operating system layer and, second,
they assume that the nodes that are going to run the system are under
the control of developers or they at least have access to them.

%Mario: Is NodIO presented here after the cut? 
% Not sure. Maybe not. - JJ
%TO DO: Add reference
There is, then, space for frameworks that do not make any of these
assumptions, such as {\sf NodIO} presented in \cite{2016arXiv160101607Manom}. {\sf NodIO} is basically 
%assumptions, such as the one presented in this paper, which will be
% named {\sf NodIO}. {\sf NodIO} is basically
% (Paloma) Is the "first" word really needed? 
% (Mario) I guess is more related to fundamentally or primarily but "basically" sounds better
a two tier client/server 
application, in which clients can run on a browser. 
Therefore, they can be embedded in any web-enabled device, including many over which 
developers have no control.

@@

%PABLO: He añadido el siguiente párrafo para el issue #31
There exist other volunteer computing frameworks written in compiled languages,
such as C++ or Java, that, in principle, can achieve better performance than
JavaScript. However, one of the advantages of our architecture is that
clients do not need to install anything in their machines, they just
need to open a web page  
% Somewhat redundant 
% Javascript is the only language that can be executed in this way without any installation, 
% also removed the default, redundant with next sentence
in a compatible browser, such as Chrome, Safari or Firefox, which are
available in almost every desktop or mobile device. Other advantage is
that the source code can be directly examined by volunteers, because it is not
compiled as other options such as Java Applets, ActionScript or
ActiveX.
Also, as we will discuss in this paper, it can provide some extra
features such as enhanced security or operating system independence.

@@

When developing volunteer based frameworks, designers must
also consider the whole {\em social} aspect in the design, with
issues related to security, trust, and privacy among others. 
The computing system becomes then a {\em techno-social system} \cite{vespignani2009predicting}.  
% (Paloma) Is the social aspect related to the word "safety"? It's not completely clear.
%a {\em social} aspect to the design of a
%computing system that makes it a {\em techno-social system}
In this paper, our intention is to
analyze the {\sf NodIO} framework from two different
%Mario: Edit this?
%present the {\sf NodIO} framework and analyze it from two different
angles: the purely technical angle, including the decisions that went
into its features, and the social angle, to account on how the persons
that voluntarily choose to enter the web page running the experiment
determine its performance and how it can be modeled and, eventually,
predicted.

The rest of the paper is organized as follows: Next we present the
state of the art (Section \ref{sec:soa}) in web-based distributed
computational systems along with attempts to predict and model its
behavior. The first results in a volunteer setting are described
in Section \ref{sec:exp1}, and experiments using a new client
architecture that uses web workers in Section \ref{sec:w2} and then
applied to a different problem in Section \ref{sec:rastrigin}. 
Finally, conclusions and future lines of work are presented in Section
\ref{sec:conclusion}.

>>>>

Using the browser as a computing resource in volunteer computing presents several advantages, 
but it remains a challenge to fully harness the browser's capabilities and
to model the user's behavior so that those capabilities can be
leveraged optimally. In this paper, we present the results of several
evolutionary computation experiments with different implementations of
a volunteer computing framework called {\sf NodIO}, designed to be
easily deployable in using free cloud resources. We use  
%Pablo: "deployable in using free cloud resources" is correct?
% why not? - JJ
different implementations to find out which one is 
able to get the user to lend more computing cycles and test different
problems to check the influence it has on said performance.
% Pablo: so the objective is "to determine which browser conf. is able
% to get more computing power from the users", it isn't? 
% (to keep this in mind the rest of the paper). What is a "browser configuration"?
% (explain it with a phrase in the abstract)
% No, we don't. Please read the paper!!!! - JJ
From these experiments we can already draw some conclusions, besides
the fact that volunteer computing can be a valuable computing
resource: the user has to be kept engaged in some way to be able to
use as many computing cycles as possible, the client has to be built
to use the computer capabilities fully, and, finally, the user
contributions follow a common statistical distribution. It is also
essential to be as open as possible with software and data.

@@

Nowadays there is a big and increasing number of distributed
evolutionary computation (EC) software frameworks available, usually
as open source frameworks \cite{Parejo12Survey}.  Most of them have
two features: they are {\em desktop} applications, that is, they must
be compiled or run directly on the operating system layer and, second,
they assume that the nodes that are going to run the system are under
the control of developers or they at least have access to them.

There is, then, space for other EC frameworks that do not make any of
these assumptions, such as {\sf NodIO} presented in
\cite{2016arXiv160101607Manom}. {\sf NodIO} implements the server of
client/server application and has some support to program an
evolutionary algorithm embedded in pages served by them which can then
communicate via that server. The client part can run in any
web-enabled device, including many over which developers have no
control, that is, volunteers.

@@

%PABLO: He añadido el siguiente párrafo para el issue #31
There exist other volunteer computing frameworks written in compiled languages,
such as C++ or Java, that, in principle, can achieve better performance than
JavaScript. However, one of the advantages of our architecture is that
clients do not need to install anything in their machines, they just
need to open a web page  
in a compatible, that is, almost any, browser, such as Chrome, Safari or Firefox, which are
available in almost every desktop or mobile device. Other advantage is
that the source code can be directly examined by volunteers, because it is not
compiled as other options such as Java Applets, ActionScript or
ActiveX are. Also, as we will discuss in this paper, it can provide some extra
features such as enhanced security or operating system
independence. This system, however, is not intended to substitute
frameworks running in dedicated hardware or the cloud, but to
complement it. In general, several volunteers will be needed to
overcome the overhead involved in using a client-server architecture
and this number of volunteers is obviously not guaranteed.

@@

One of the aspects that must be taken into account in volunteer based
frameworks is its {\em social} aspect in the design, with 
issues related to security, trust, and privacy among others. 
The computing system becomes then a {\em techno-social system}
\cite{vespignani2009predicting}. In that sense, its social aspects
cannot be separated from its technical specifications, with one
feeding into the other. One of these aspects is safety for the user
and the scientist running it: the user must be reasonably sure that the
experiment is not misusing its resources and the scientist must have a
certain safety in the sense that its server is not going to be
attacked via cheating or denial or service attacks. We have decided to
approach this in a pragmatic way. Including many security features
would decrease performance; on the other hand, we publish all data
resulting from the experiment once private information, such as IPs,
has been processed to a untraceable hash. The experiments themselves
are a proof that there are enough users trusting the experiment to
give it its CPU time and that there has been no attempt to cheat by
sending specially crafted chromosomes or to attack the server to make
it stop working because there was no incentive in doing so. The
system, besides, is running in a dedicated Platform as a Service
container or Virtual Private Server so accessing the server would only
show a few data obtained from the experiment, nothing more. Making it
free software ensured this was known by anyone in the experiment, via
links in the web page that hosted it.

The rest of the paper is organized as follows: Next we present the
state of the art in web-based distributed
computational systems along with attempts to predict and model its
behavior  in Section \ref{sec:soa}. The first results in a volunteer
setting are described 
in Section \ref{sec:exp1}, and experiments using a new client
architecture that uses web workers in Section \ref{sec:w2} and then
applied to a different problem in Section \ref{sec:rastrigin}. 
Finally, conclusions and future lines of work are presented in Section
\ref{sec:conclusion}.
